\textbf{\sc Towards Deliberative Control of Autonomous Robots}

### 2012-02-09 arch.tex
updates

The \rx architecture was designed with the objective to bring planning
closer to the low-level control of the vehicle while maintaining its
reactivity. Indeed, in most of the architecture where task planning is
embedded, this task remains at a failry high level of abstraction as
the complexity of this task would otherwise impede the system
reactiveness. As a result the planning problem managed {\em in situ}
remains fairly detached from low-level control and most of the
reactive adaptation which in contrast is managed by specialized
component that relies on different representations for modeling the
local behavior they control or even do not have any explicit formal
model of this behavior. The two angles our architecture attack this issue 
are the following:

@@

\item Distribute the global planning problem into multiple decision
  loops, called {\em reactors}. Each reactor have its own scope both 
  functionally (one for example focusing on the conversion of high 
  level waypoints into low-level commands) and temporally as 
  each reactor specifies to the agent their planning look-ahead and
  their expected latency before being able to produce plan.

\item Enforce a stronger interaction between the planning process 
  and the integration of the world evolution. The objective being to
  make sure that the plannign task is done while continuously informed 
  on new world state change. By doing so, we expect to allow the planner
  tohave better informed decision despite the fact that its latency is
  larger that the world change rate.

@@

For the first point \rx provides a formal framework described in
details in \cite{py10} and \cite{rajan12} and is not the focus of this
paper. The second have deep implication on how on should integrate a
planner in this framework in general and was applied to the europa
framework specifically. We'll develop later how this requirement
implies to see in-situ planning as two parallels processes that are
tightly linked through the same plan structure. But before that we
need to introduce at least the high leve lconcepts and ideas behind
our architecture as a whole.

\rx architecture is structured around the notion of {\em state
  variables} as the unique basis of interactionalong with a well
defined ownership model for each of these.  In this framework each
reactor is seen as a black box providing its own {\em internal} state
variables for which its is reponsible to maintain its value at every
single instant -- represented by a {\em tick} with a fixed
duration. In order to identify its state a reactor can subscribe to
{\em external} state variables for which it will received new updates
when the owner of these will produce its {\em internal} state. This
{\em synchronization} process progate through the reactor dependency
graph on a bottom up flow at every {\em tick} ensuring that all the
reactors have a consistent view of the world. Conversely each reactor
can request a future change on one of its {\em external} state
variable which will be transfromed as an {\em internal} goal to the
owner. This goal is expected to porduce a deliberation phase on the
receiver which can in turn produce a plan impacting its {\em external}
state variables that will propagate down following the same mechanism.

@@

Take for example the \rx agent instance we use in our AUV
(Fig. \ref{fig:agent}). In this architecture all the reactors
(symbolized by colored boxes) do not have explicit connection from one
to another but instead rely on a publish/subscribe model on state
variables. For Example the \textsf{Pilot} reactor needs the {\em
  externally} managed state informations from \textsf{Sensor\_Data}
and \textsf{Nav\_cmd} states variables in order to maintain {\em
  internally} the state variables \textsf{Waypoint} and
\textsf{Nav\_policy}. This relation applies in two way :

@@

The way the agent interleave planning \texttt{step}s and
synchronization for a single reactor is illustrated in Figure 
\ref{fig:tick-exec}. Synchronization happens at every single 
tick and will interrupt planning allowing this reactor to identify
 its current state that can propagate through the reactor graph
 in order to ensure t aconsitent view of the state of the world 
between all the reactors.

@@

The \eu reactor is our implementation inside \rx of a reactor that
usess automated planning for its execution. This reactor is fully
based on the europa framework and applies the core idea of \rx we
described above to implement a reactor for which the sole purpose is to
deliberate as time advance. The design of this reactor takes cues from 
the IDEA architecture presented on \cite{mus02, mus06} in the sense taht both
planning and execution are based on a single unified model. The choice 
of \eu as the planning framework was directed by the fact that this 
framework provide a rich and highly configurable toolset to be able to 
infer around a rich temporal domain.

@@

Now, consider that at the next tick $\tau+1$ the reactor managing the
agent location was able to change the location state varaibale to the
\texttt{Going}. \rx informs our reactor of this new observation which
is added in the plan structure as a fact -- starting exactly at
$\tau+1$ and lastign for an unlknown duration which is at least 1
tick. As this new observation is integrated in the plan through
synchronization the solver can identify the similarity between this
new observation oand the next token in the existing plan. Which mean
that instead of making a blind exhaustive search of all the possible
implication the synchronization process starts its execution with the
assumption that these new observations are consitent with the current
plan maintained. As a result at the end of synchronization we will
have the plan depicted in Fig. \ref{fig:shop:exec1}. Which just
propagated the information provided by the new observation in the
exisiting plan (restricting in turn the start time of \texttt{Going}
and by propagtion of the constraints \texttt{Go} to be {\em exactly}
$\tau+1$). While focusing uniquely on the execution frontier ($\tau+1$
in this figure), the europa solver took benefit of the exisiting plan
to evaluate a solution that asumes that the observation is consistent
with what was decided inside this reactor.

>>>>

A prime motivation behind the \rx architecture was to design a
controller which could bring automated planning closer to low-level
control of the vehicle while maintaining its reactivity. Traditionally
planning has been deemed computationally expensive with planner
performance restricting robot reactivity
\cite{ghallab04,Dias:2003ua}. In control architectures where task
planning is embedded, planning consequently remains at an abstract
level with the assumption that it would otherwise impede system
reactiveness and that the environment can change at a faster rate than
the planner can plan for it. In such a situation, the agent may thrash
if the internal state of the plan gets out of synch with the actual
state of the world. As a result the planning problem managed {\em in
  situ} remains fairly detached from low-level control. Most
adaptations at the lower-level reactive layers in contrast are managed
by specialized components (for example GeNoM \comment{} and CLARATy
\comment{}) that rely on different representations for modeling local
control behavior or worse do not have any explicit formal agent
behavior model. The consequent different techniques for specifying
each layer in the architecture results in duplication of effort and a
diffusion of knowledge leading to significant design and integration
issues \cite{ds1report}.  \texttt{IDEA} mitigated most of these issues
by tightly integrating planning and execution in a single
representational and computational framework with a unified
declarative model. Our approach with \texttt{T-REX} builds on this
design methodology albeit in a substantially systematic manner

The \rx architecture contributes to agent architectures over and
beyond \texttt{IDEA} are in the following ways:

@@

\item \emph{Partitioning} the global planning problem into multiple
  decision loops, called {\em reactors}. Each reactor has its own
  scope both functionally (one for example focusing on the conversion
  of high level waypoints into low-level commands) and temporally as
  each reactor declares its planning look-ahead and expected latency
  before being able to produce plan to the agent depending on where it
  is located in a reactor dependency graph such as shown in
  Fig. \ref{fig:agent}.

\item \emph{Coupling} a tighter integration between the planning
  process and world evolution to ensure that planning occurs while
  being continuously informed of state change. By doing so, the
  planner is better informed in its deliberation despite a larger
  latency than changing world change state.

@@

\rx provides a formal framework described which details partitioning
in \cite{py10} and \cite{rajan12} and is not the focus of this
chapter. Coupled interaction between planning and the world on the
other hand, has a significant implication on how planning can be
integrated in such frameworks; here we show using the \eu planner can
be coupled to our framework. First we link concepts in \eu planning
with the \rx framework.

% on should integrate a
% planner in this framework in general and was applied to the europa
% framework specifically. We'll develop later how this requirement
% implies to see in-situ planning as two parallels processes that are
% tightly linked through the same plan structure. But before that we
% need to introduce at least the high leve lconcepts and ideas behind
% our architecture as a whole.

The \rx architecture is structured around the notion of \eu {\em state
  variables} as the fundamental basis of interaction. Additionally it
comes with a well-defined ownership model associated with each state
variable. In this framework each reactor is seen as a 'black box'
providing its own {\em internal} state variables which it is
responsible for in order to maintain its value at every instant of
time. Each such instance is represented by a {\em tick} with a fixed
duration; in our implementation of \rx on the Dorado, we use a tick of
$1$ second. In order to identify its state a reactor can subscribe to
{\em external} state variables owned by other reactors, for which it
will receive new updates when the owner reactor generates its {\em
  internal} state.

This {\em synchronization} process propagates up through the reactor
dependency graph bottom up at every {\em tick} ensuring that all the
reactors have a consistent view of the world. Conversely each reactor
can request a future change on one of its {\em external} state
variables which will be transfromed as an {\em internal} goal to the
owner. Such \emph{goal dispatching} is expected to produce a
deliberation phase for the receiving reactor which can cyclically and
in turn, produce a plan impacting its {\em external} state variables
that will propagate down following the same mechanism.

@@

Consider for example the \rx agent instance we use on our AUV as shown
in Fig. \ref{fig:agent}. In this instance all the reactors (symbolized
by colored boxes) do not have an explicit connection from one to
another but instead rely on a publish/subscribe model on shared state
variables. For instance the \textsf{Pilot} reactor needs the {\em
  externally} managed state information from \textsf{Sensor\_Data} and
\textsf{Nav\_cmd} state variables in order to maintain the internal
state variables \textsf{Waypoint} and \textsf{Nav\_policy}. This
relation applies in two ways:

@@

The way the agent interleave planning \texttt{step}s and
synchronization for a single reactor is illustrated in
Fig.~\ref{fig:tick-exec}. Synchronization happens at every single tick
and will interrupt planning allowing this reactor to identify its
current state that can propagate through the reactor graph in order to
ensure t aconsitent view of the state of the world between all the
reactors.

@@

The \eu reactor is our implementation inside \rx of a reactor that
usess automated planning for its execution. This reactor is fully
based on the europa framework and applies the core idea of \rx we
described above to implement a reactor for which the sole purpose is
to deliberate as time advance. The design of this reactor takes cues
from the IDEA architecture presented on \cite{mus02, mus06} in the
sense that both planning and execution are based on a single unified
model. The choice of \eu as the planning framework was directed by the
fact that this framework provide a rich and highly configurable
toolset to be able to infer around a rich temporal domain.

@@

Now, consider that at the next tick $\tau+1$ the reactor managing the
agent location was able to change the location state varaibale to the
\texttt{Going}. \rx informs our reactor of this new observation which
is added in the plan structure as a fact -- starting exactly at
$\tau+1$ and lastign for an unlknown duration which is at least 1
tick. As this new observation is integrated in the plan through
synchronization the solver can identify the similarity between this
new observation oand the next token in the existing plan. Which mean
that instead of making a blind exhaustive search of all the possible
implication the synchronization process starts its execution with the
assumption that these new observations are consitent with the current
plan maintained. As a result at the end of synchronization we will
have the plan depicted in Fig. \ref{fig:shop:exec1}. Which just
propagated the information provided by the new observation in the
exisiting plan (restricting in turn the start time of \texttt{Going}
and by propagtion of the constraints \texttt{Go} to be {\em exactly}
$\tau+1$). While focusing uniquely on the execution frontier ($\tau+1$
in Fig. \ref{fig:shop:exec1}), the europa solver took benefit of the
exisiting plan to evaluate a solution that asumes that the observation
is consistent with what was decided inside this reactor.
